\bf Baseline Drift Estimation for Time Series Data Using Quantile Trend Filtering

### 2019-03-18 Manuscript/detrendify_JCGS.tex

The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest. As the size of the data and the number of quantiles grows, all solvers will eventually break \Halley{Eric}{Can we be more specific about what we mean when we say ``break?"}. To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously. We propose a divide-and-conquer approach via an ADMM algorithm for solving large problems in a piecewise fashion. The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is  described in greater detail by \cite{boyd2011distributed}, but we briefly review how it can be used to iteratively solve the following equality constrained optimization problem.
\begin{equation}
\label{eq:split_objective}
\begin{split}
&\text{minimize} \; f(\V{\theta}) + g(\Vtilde{\theta}) \\
&\text{subject to} \; \M{A}\V{\theta} + \M{B}\Vtilde{\theta} = \V{c}.
\end{split}
\end{equation}
\Eric{Joe}{I'm already lost here. Can we give some context to what $\theta$ and $\widetilde{\theta}$ are in the context of our windowing algorithm? Maybe we should say something about the windowing first before we dive head first into Lagrangians? What are $A$ and $B$? Do these simplify to something in our specific application? The stuff below might seem obvious to you, but I'm totally lost. It might be better to talk about the notation for the windows first, and how different windows will give different thetas for the overlapping parts, and then talk about equality constrained optimization as a solution for the overlapping parts.} Recall that finding the minimizer to an equality constrained optimization problem is equivalent to the identifying the saddle point of the Lagrangian function associated with the problem \Eqn{split_objective}. ADMM seeks the saddle point of a related function called the augmented Lagrangian,
%The Lagrangian for the ALM problem
\begin{eqnarray*}
\mathcal{L}_{\gamma}(\V{\theta},\Vtilde{\theta},\V{\omega}) & = & f(\V{\theta}) + g(\Vtilde{\theta}) + \langle \V{\omega}, \V{c} - \M{A}\V{\theta} - \M{B}\Vtilde{\theta} \rangle
+ \frac{\gamma}{2} \lVert \V{c} - \M{A}\V{\theta} - \M{B}\Vtilde{\theta} \rVert_2^2,
\end{eqnarray*}
where the dual variable $\V{\omega}$ is a vector of Lagrange multipliers and $\gamma$ is a nonnegative tuning parameter. When $\gamma = 0$, the augmented Lagrangian coincides with the ordinary Lagrangian.

@@

	\cite{chen2008} proposed the extended Bayesian Information Criteria (eBIC), specifically designed for large parameter spaces.
	\begin{eqnarray*}
	\label{eq:eBIC}
	\mbox{eBIC}_{\gamma}(s) & = & -2\log(L\{\Vhat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
	\end{eqnarray*}
	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given model. \Halley{Eric}{What is $s$?} We used this criteria with $\gamma = 1$, and $P=n-k-1$. \Fig{BIC} illustrates the difference among the scaled, unscaled ($\sigma = 1$), and scaled extended BIC criteria
	when applied to a single dataset replicate from our simulation study. In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.

>>>>

The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest. As the size of the data and the number of quantiles grows, eventually the matrix of constraints will no longer fit into memory. To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously. We propose a divide-and-conquer approach via an ADMM algorithm for solving large problems in a piecewise fashion. The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is  described in greater detail by \cite{boyd2011distributed}, but we briefly review how it can be used to iteratively solve the following equality constrained optimization problem.

\begin{equation}
\label{eq:split_objective}
\begin{split}
&\text{minimize} \; f(\V{\theta}) + g(\Vtilde{\theta}) \\
&\text{subject to} \; \M{A}\V{\theta} + \M{B}\Vtilde{\theta} = \V{c}.
\end{split}
\end{equation}
\Eric{Joe}{I'm already lost here. Can we give some context to what $\theta$ and $\widetilde{\theta}$ are in the context of our windowing algorithm? Maybe we should say something about the windowing first before we dive head first into Lagrangians? What are $A$ and $B$? Do these simplify to something in our specific application? The stuff below might seem obvious to you, but I'm totally lost. It might be better to talk about the notation for the windows first, and how different windows will give different thetas for the overlapping parts, and then talk about equality constrained optimization as a solution for the overlapping parts.} Recall that finding the minimizer to an equality constrained optimization problem is equivalent to the identifying the saddle point of the Lagrangian function associated with the problem \Eqn{split_objective}. ADMM seeks the saddle point of a related function called the augmented Lagrangian,
%The Lagrangian for the ALM problem
\begin{eqnarray*}
\mathcal{L}_{\gamma}(\V{\theta},\Vtilde{\theta},\V{\omega}) & = & f(\V{\theta}) + g(\Vtilde{\theta}) + \langle \V{\omega}, \V{c} - \M{A}\V{\theta} - \M{B}\Vtilde{\theta} \rangle
+ \frac{\gamma}{2} \lVert \V{c} - \M{A}\V{\theta} - \M{B}\Vtilde{\theta} \rVert_2^2,
\end{eqnarray*}
where the dual variable $\V{\omega}$ is a vector of Lagrange multipliers and $\gamma$ is a nonnegative tuning parameter. When $\gamma = 0$, the augmented Lagrangian coincides with the ordinary Lagrangian.

@@

	\cite{chen2008} proposed the extended Bayesian Information Criteria (eBIC), specifically designed for large parameter spaces.
	\begin{eqnarray*}
	\label{eq:eBIC}
	\mbox{eBIC}_{\gamma}(\nu) & = & -2\log(L\{\Vhat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
	\end{eqnarray*}

	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod. We used this criteria with $\gamma = 1$, and $P=n-k-1$. \Fig{BIC} illustrates the difference among the scaled, unscaled ($\sigma = 1$), and scaled extended BIC criteria
	when applied to a single dataset replicate from our simulation study. In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
